Exploring environmental impacts of colonial contacts on Indigenous communities in northeastern Taiwan

### 2024-07-26 analysis/paper/paper.qmd
fix the render issue and edit

---
title: "Exploring environmental impacts of colonial contacts on Indigenous communities in northeastern Taiwan"
author:
  - Li-Ying Wang:
      correspondence: "yes"
      email: liyingw15@gmail.com
      orcid: 0000-0001-9060-2125
      institute:
        - AS
institute:
  - AS:
      name: Academia Sinica
      address: 
  - IHP: 
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

# Introduction
<!-- introduction about European expansion -->
European expansion since the 15th century marked the beginning of an era characterized by global trade, exploration, and colonization, which often led to profound impacts on indigenous societies in many parts of the world. The forms of European colonization and power relations varied across regions, depending on specific spatial and temporal contexts [@Jorden2009]. Understanding the dynamics of these colonial processes has become a significant issue in archaeology, as it provides critical insights into the complex outcomes of European colonization [@Gosden2004]. In Asia, European colonization was primarily driven by the procurement of Asian luxury goods, raw materials, or natural resources for commercial profit. The new exchange systems and governance policies imposed by European colonizers often had profound effects on local economic systems, social structures, and settlement patterns [@Christie1996; @Junker2004; @Lape2002]. However, Indigenous communities did not just passively accept European colonization. Recent discussions on colonial interactions also highlight the diverse interactions employed by Indigenous communities in response to European presence, such as resistance, negotiation, persistence, resilience, and entanglement [@Torrence2003; @Silliman2005; @Mitchell2010; @Jorden2009; @Alexander1998].

<!-- NE Taiwan as a case study -->
While many studies focus on prominent colonies, places on the periphery of colonial center were also more and less impacted by European powers or associated trading systems [@Acabado2017; @Trabert2017]. Exploring these areas is important to gain a broader understanding of the varied consequences of colonial expansion. Northeastern Taiwan, specifically Yilan, serves as an example of a peripheral case in the 17th century when Europeans established their colonial base in northern Taiwan. Due to its geologically challenging and relatively inaccessible location, Yilan experienced limited direct European control. However, local settlements were frequently involved in the regional exchange network stimulated by the colonial presence in the north. In the mid-17th century, the VOC (Dutch East India Company) formalized the deerskin trade in Taiwan for export in a large scale, involving many Indigenous communities and Han Chinese settlers [@Heyns2001]. Despite some discussions on the policies and operations of the deerskin trade, direct investigations on archaeological evidence into the experiences of local indigenous people to the commercial activities are rare. The European impacts on local ecological conditions have been recorded in many places in colonial settings [@Jorden2009]. Whether the interaction between Indigenous people and their environment changed due to European trade demands requires further archaeological investigation.

@@

# Colonial encounter in Northeastern Taiwan
<!-- the structure of power relations, colonial control in particular spatial and temporal contexts-->
In 1626, the Spanish colonized northern Taiwan and built fortress in Tamsui and Keelung to the threat of the VOC (Vereenigde Oost-Indische Compagnie; Dutch East India Company) in Asian market, which established a colony in southern Taiwan two years before the Spanish in the north. In 1642, the Dutch defeated the Spanish and occupied their fortresses, officially expanding their colonial power to northern Taiwan until the end of their colonization in Taiwan in 1662 [@Andrade2007]. The form of European colonization in northern Taiwan relied to a certain degree on local Indigenous people for the development of the colony and resources collections among different Indigenous groups. For example, the Basay people, an indigenous group active in Keelung, were recorded to have assisted in the colonial activities of the VOC administrators, such as coordinating between indigenous villages [@Kang2016]. The colonial domination was mostly established through the conquest over indigenous villages, but the colonial control varied depending on the location of indigenous settlements. For most indigenous villages distant from the European colonial base, the colonial control was relatively less, indirect, and even unsuccessful.
<!-- domination and the processes by which it was established and resisted, how and why colonial powers were able to exert control (however limited) over Native populations -->

Kavalan, or the present day Yilan, in northeastern Taiwan is an example of slight European presence for most of the time. Kavalan is about 70 kilometers away from the European northern fortress with mountains in between. Local indigenous groups were Kavalan people. According to the Dutch census, there were more than 40 Kavalan villages with approximately 10,000 people in the 17th century [@Nakamura1938], which were also mentioned by the Chinese and Japanese in later periods, suggesting a continuous indigenous habitation until the 19th century. Kavalan people first experienced European powers in 1632, when they were attacked by Spanish traders [@citation]. Later in 1647, a group of Dutch soldiers attacked several Kavalan villages and asked those defeated villages to pay tributes [@Kang2016]. According to Dutch written record, the VOC had built a trading outpost in Trobiawan, located near the coastal area in Yilan, but there might be not much officials at any given time. It appears that the Kavalan area was in the situation where trading activities were frequent and active stimulated by the European presence in northern Taiwan, with occasional pressures from colonial powers. The European control over the Kavalan must have been largely indirect and limited, which allowed the maintenance of autonomy and flexibility of Kavalan people for negotiation tributes requested by the VOC, such as deerskins or products [@Kang2016].

Before the arrival of the Europeans in northern Taiwan, there had been a regional exchange network involving Chinese merchants, Japanese traders, and indigenous groups in northern and northeastern Taiwan since the 14th century [@citation]. The local indigenous groups traded

The Yilan Plain is the major coastal plain in northeastern Taiwan. Prehistory in the Yilan Plain begins from the Middle Neolithic period around 4000 B.P. and continues throughout the Metal Age period (2000-400 B.P.) (Liu 2011). The final Late Iron Age in Taiwan can be viewed as a protohistoric period due to the interaction of record-keeping Chinese and Europeans with local cultures. Most Late Metal Age sites in the Yilan Plain can be related to the local Indigenous people, the Kavalan, according to oral history and similarities between historical Kavalan material culture and archaeological artifacts (Chen 2004).

 Yilan is about 70 kilometers away from the first European northern base with mountains in between. Indigenous people in Yilan were not in direct contact with colonial powers until 1632, when they were attacked by the Spanish. Later in 1647, the Dutch attacked the Indigenous people and forced them to accept colonial rules and economic demands by paying annual tribute (Andrade, 2007). The Dutch also built a trading outpost in Trobiawan. According to the Dutch census, Indigenous people in northeastern Taiwan occupied over 40 settlements with approximately 10,000 people in the 17th century (Nakamura 1938), and those settlements were also described by the Chinese in later periods, demonstrating a continuous Indigenous habitation until the 19th century (Figure 1).
Due to the fear of European attacks, most local communities in Yilan maintained good relationships with the Europeans by paying tributes and Indigenous leaders could obtain foreign materials as gifts according to the degree of allegiance (Kang 2012). In addition to historical records, archaeological evidence also shows complex inter- and intra- exchange networks based on the findings of foreign goods, such as stonewares commonly found in European outposts in Southeast Asia, Chinese porcelains, and glass and stone beads (Chen 2005; Chen 2007; Liu 2011; National Museum of Taiwan History 2005, Wang and Liu 2007). This suggests frequent trade activities since the protohistoric period in the 14th century and later intensified through the global trade network stimulated by the Europeans and associated colonial activities. Moreover, foreign materials such as ceramics, beads, and textile were viewed as prestige goods in local communities before European arrival (Li and Wu, 2006).

Northeastern Taiwan is an especially interesting example of European and Indigenous interaction because the impact of European colonialism appears to have been much less pronounced compared to other places in this region. I define northeastern Taiwan as an area at the periphery of colonial center where the European power was indirect with a few sporadic, mostly non-violent direct contacts. This type of context has also been called pericolonialism (Acabado, 2017; Kang, 2016, pp. 31-78; Trabert, 2017). My PhD study at Kiwulan (14-19th A.D.), the biggest Indigenous settlement during the 17th century in Yilan, indicates that European contact and foreign goods resulted in an increased social differentiation in the local society through analyses of trade items, pottery production, and burial goods (Wang and Marwick, 2020a, 2020b, 2021). The results demonstrate that there was an indirect effect of European colonization in this region that could lead to changes in the organizational mode of a society that have not been identified before.

In addition to social changes, how did the interactions with Europeans and Chinese lead to changes in the local ecological system need further examination. For example, during the Dutch colonial period, deer hide was especially demanded in the Dutch colonization period as payment for tributes due to its high value for export to Japan, which led to intensified deer hunting in historical Taiwan (Andrade, 2007; Koo, 2011). While the commercial activities and a substantial decrease in Taiwan’s deer population were well documented, much less is known about how this historical consequence can be situated in a local ecological framework. Did the participation in these trade activities result in a change in Indigenous subsistence behaviors or the human-environment interactions? What was the response of local Indigenous people to the decrease of deer population? Those questions are important because they can help us to trace back the Indigenous cultures and their resilience under colonial interference since the protohistoric Taiwan.

## trade and exchange of Han Chinese

In this research, I will explore whether and how the introduction of foreign materials and practices, new forms of production, and migrated people from Europe and Chinese contributed to the changes in Indigenous social, environmental, and ecological systems since the protohistoric period (Liu and Wang, 2017). I will focus on archeological evidence from different parts in northeastern Taiwan to compare Indigenous reactions in a pericolonial context (Acabodo, 2017). Yilan is an ideal location as a starting point because of the frequent foreign contact with Europeans and Chinese traders in the protohistoric and historic period. Also, the ecological variation across the landscape from the coastal area to mountains provides multiscalar variability for exploring the effect of colonial contact on a regional scale. As a place located in the surrounding area of the European colonial base in northern Taiwan, Yilan enables a pericolonial perspective to understand the effect of European colonization. I will use historical ecological approaches to examine the interactions of people and environment in a colonial context, and their consequences on the development of the past and contemporary culture in a long-term view (Braje, 2017). This project will help us to understand social and ecological reactions of Indigenous communities to European colonization and cross-cultural contact that has not yet been fully understood.

I will use historical ecology as the conceptual framework to understand the human-environment interactions in northeastern Taiwan, which holds that human impacts on the ecological system are historically dependent (Balée, 1998; Braje, 2017; Fitzhugh et al., 2019). In a colonial context, historical ecological investigations can help us to understand the impacts of European presence and associated economic activities on the local environments and domestic economies (Oas and Hauser, 2018; Wallman et al., 2018). Located in northeastern Taiwan, Yilan is close to the Dutch colonial base in northern Taiwan with favorable soil conditions for rice growing and natural resources such as forests and deer. The frequent exchange activities with Europeans and Chinese merchants, and local interests in foreign goods could lead to further exploitation of deer and intensified rice growing as tributes for maintaining good relations with Europeans and exchanging more imported goods. 
My hypothesis is that if European colonial activities resulted in environmental and ecological changes in northeastern Taiwan due to deer hunting and rice growing, then we could observe changes in Indigenous subsistence and land use patterns. I will test my hypothesis by studying archaeological sediments, animal and plant remains, organic residues in locally made ceramics, and site locational preferences of Late Metal Age sites in Yilan to identify any changes after the colonial presence.

# Methods
The later period of Chinese migration will also be included to provide a long-term perspective

## Geoarchaeological and geochemical analysis for sediments
To explore changes in environments, I will examine sediments from archaeological sites in northeastern Taiwan, Yilan. I will mainly focus on archaeological sites recorded as Indigenous settlements that can be traced back to the protohistoric and historic period when they experienced the European and Chinese presence, such as Kiwulan I, Kiwulan II, and Yilan Agricultural Vocational High School site. Kiwulan and Yilan Agricultural Vocational High School are two major Late Metal Age sites with extensive excavation in recent years. Kiwulan (600-100 B.P.) is located at the northern part of Yilan plain (Chen, 2007), while Yilan Agricultural Vocational High School site (400-200 B.P.) is situated on the central part of Yilan (Li, 2014). Archaeological evidence from both sites includes a large amount of pottery sherds, imported stoneware and porcelains, wooden artifacts, stone and metal tools, pipes, thousands of glass beads, agate beads and metal ornaments, post holes with in-situ wooden posts, and burials. Based on their locations and the historical records, those two sites can be traced back to the 17th century when they were named Banouran and Beili according to the Dutch census (Nakamura 1938). Those archaeological materials are mainly stored in the Cultural Affairs Bureau in Yilan County, and it is feasible to analyze those materials and collect data based on previous collaboration during my PhD. 
For archaeological sediments, I will use geoarchaeological methods to measure the physical and chemical properties of sediments, including particle size, magnetic susceptibility, pH values, electrical conductivity, and organic matter and carbonates proportions to understand soil conditions and Indigenous landuse. I will also use organic chemical analysis to detect lipids in sediments to reconstruct past environments, fire and land use patterns associated with the colonial presence and long-term cultural interaction. Geoarchaeological and geochemical methods are useful and insightful methods especially when faunal and botanical evidence were poorly preserved, such as in a condition of acidic environments.

## Faunal and botanical analysis, and organic residue analysis
To explore any changes in the subsistence strategies, I will examine faunal and botanical evidence directly if they were preserved in good conditions. The historical records point out frequent exchange activities between Europeans, Chinese traders, and different Indigenous people, such as Kavalan and Basay during this period (Chen, 2008). According to historical records, Indigenous people in Yilan exchanged natural resources, such as deer hide and rice, for porcelains, metal items, and beads from Europeans, Chinese traders, and other Indigenous people, such as Basay (Borao, 2001). Deer hide was especially demanded in the Dutch colonization period as payment for tributes due to its high value as a commodity in East Asia. 
To understand change in subsistence behaviors or the human-environment interactions during and after the colonial period, I will examine two types of evidence from those sites. First is zooarchaeological and botanical data related to economic activities in a colonial context, especially animal bones such as deer and plant remains such as rice grains. Second, I will build on my previous pilot studies of isotope geochemical analysis for samples from Kiwulan, and greatly expand the sample sizes to include human remains (e.g., teeth), animal bones, and ceramic fabrics. This will enable me to measure dietary or subsistence changes related to environments, and stability before and after the European presence, and Chinese presence. 
To accomplish those analyses, I will draw on my existing collaborations with the Paleoclimatology and Organic Geochemistry lab in the University of Washington and the Microbial Metabolic Diversity lab in the Academia Sinica for organic residue analysis and DNA analysis, especially for pottery sherds. By analyzing the composition of the organic residues absorbed and trapped in potsherds, we can know what categories of food are cooked in pots, such as animal fats, plants, and aquatic resources that provides direct evidence to study ancient diet (Evershed 2008; Craig 2011; Cramp 2011). Similarly, microbes from food residues in pots offer a way to detect ancient diet because different types of microbes may be associated with different types of food (Bokulich et al. 2016). I will combine those advanced techniques to understand subsistence strategies in northeastern Taiwan.

>>>>

---
title: "Exploring environmental impacts of colonial contacts on Indigenous communities in northeastern Taiwan"
author:
  - Li-Ying Wang:
      correspondence: "yes"
      email: liyingw15@gmail.com
      orcid: 0000-0001-9060-2125
      institute: Academia Sinica
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

# Introduction
<!-- introduction about European expansion -->
European expansion since the 15th century marked the beginning of an era characterized by global trade, exploration, and colonization, which often led to profound impacts on indigenous societies in many parts of the world. The forms of European colonization and power relations varied across regions, depending on specific spatial and temporal contexts [@Jordan2009]. Understanding the dynamics of these colonial processes has become a significant issue in archaeology, as it provides critical insights into the complex outcomes of European colonization [@Gosden2004]. In Asia, European colonization was primarily driven by the procurement of Asian luxury goods, raw materials, or natural resources based on commercial interest [@Prakash1997]. The new exchange systems and governance policies imposed by European colonizers often had profound effects on local economic systems, social structures, and settlement patterns [@Christie1996; @Junker2004; @Lape2002]. However, Indigenous communities did not just passively accept European colonization. Recent discussions on colonial interactions also highlight the diverse interactions employed by Indigenous communities in response to European presence, such as resistance, negotiation, persistence, resilience, and entanglement [@Torrence2003; @Silliman2005; @Mitchell2010; @Jordan2009; @Alexander1998].

<!-- NE Taiwan as a case study -->
While many studies focus on prominent colonies, places on the periphery of colonial center were also more and less impacted by European powers or associated trading systems [@Acabado2017; @Trabert2017]. Exploring these areas is important to gain a broader understanding of the varied consequences of colonial expansion. Northeastern Taiwan, specifically Yilan, serves as an example of a peripheral case in the 17th century when Europeans established their colonial base in northern Taiwan for a better access to Asian market. Due to its geologically challenging and relatively inaccessible location, Yilan experienced limited direct European control. However, local settlements were frequently involved in the regional exchange network stimulated by the colonial presence in the north. In the mid-17th century, the VOC (Dutch East India Company) formalized the deerskin trade in Taiwan for export in a large scale, involving many Indigenous communities and Han Chinese settlers [@Heyns2001; @Cheng2017]. Despite some discussions on the policies and operations of the deerskin trade, direct investigations on archaeological evidence into the experiences of local indigenous people to the commercial activities are rare. The European impacts on local ecological conditions have been recorded in many places in colonial settings [@Jorden2009]. Whether the interaction between Indigenous people and their environment changed due to European trade demands requires further archaeological investigation.

@@

# European encounters in the northern region of Taiwan
<!-- the Spanish colonization and the power relations in particular spatial and temporal contexts-->
In 1626, the Spanish East Indies Company (Indias Orientales Españolas) arrived in northern Taiwan and established the San Salvador colony at Heping Dao in Keelung, followed by Santo Domingo in Tamsui, to counter the threat of the VOC (Vereenigde Oost-Indische Compagnie; Dutch East India Company) in the Asian market, which had established a colony in southern Taiwan two years earlier [@Borao2009, p.41; @Andrade2007]. Despite the intention to establish a colonial city, San Salvador remained in its early developmental stages, primarily consisting of a fortress and a church, with a population of Spanish officers, soldiers, and missionaries [@Borao2009, pp. 103-134; @Berrocal2018]. The area neighboring San Salvador was inhabited by the Taparris and Quimaurris, local indigenous groups who had resided at the location of the fort before being expelled by the Spanish [@Borao2001, SIT:165, p.81]. During Spanish colonization (1626-1642), the colonial processes may not have involved extensive colonialism or asymmetrical power dynamics, with territorial control likely limited to the fort. Written records mention that the Spanish appeared to rely on the local indigenous people for exchanging local goods and food supplies [@Borao2001, SIT:170-171]. This is further supported by archaeological evidence at San Salvador [@Berrocal2018], where only a few European materials were found, alongside more indigenous and Chinese items that might have been used by the Europeans, indicating cultural entanglement.

<!-- the Dutch domination and the processes by which it was established, how and why colonial powers were able to exert control -->
In 1642, the Dutch defeated the Spanish and took over the San Salvador and the Santo Domingo, expanding their colonial power to northern and eastern Taiwan until 1662. Compared to the Spanish, the Dutch were more strategic in consolidating their power by forming alliances with indigenous villages, implementing political and economic institutions, and conquering villages that were unwilling to comply [@Kang2016; @Andrade2007]. For example, the VOC governors used various colonial institutions to secure their authority, including annual landdag gathering, taxation, tribute system, village lease system, census investigations, and the selection of indigenous representatives [@Kang2016; @Andrade1997; @Andrade2007]. Among the political institutions, the tribute system operated from 1642 to 1647 and village lease system introduced in 1644 had great impacts on local indigenous economy and environment. Based on tribute system, each household of allied villages was required to pay rice or high-quality deerskin annually. The amount of deerskins collected from indigenous people in northern Taiwan once reached 7,000 pieces in 1646 [@Andrade2007]. Moreover, to increase deerskin exports to Japan, the Dutch implemented a hunting license system (1636-1645) that gave Han Chinese permission to hunt deer on indigenous lands during certain periods [@Heyns2001]. From 1633 to 1661, the average of deerskins exported from Taiwan was about 67,000 per year [@Koo2011; @Cheng2017]. Those colonial activities accelerated the decline in the deer population and could lead to a harsh situation for the indigenous economy .

<!--the variation of colonial powers over native populations, indirect colonization at Kavalan-->
Although Dutch power expanded across nearly all of Taiwan in the mid-17th century, their colonial control and domination varied significantly in different geographical regions. In areas distant from the European colonial bases, colonial control was often challenging and largely unsuccessful. Yilan in northeastern Taiwan is an example of minimal European presence and indirect colonization, with fewer imposed colonial rules for most of the time [@Kang2012]. Yilan is an alluvial plain located about 70 kilometers from the European colony in Keelung, separated by mountains, and inhabited mainly by the local indigenous group, the Kavalan. Given its strategic location on the route to Torrobouan, where the Dutch believed gold mines existed, they were eager to bring this area under their control [@DZI2000; @Kang2016, p47]. In 1644, the Dutch invaded Yilan to incorporate the defeated Kavalan villages into their northern colonial territory, and demanded tribute from the indigenous people [@Kang2016]. nstead of engaging directly with the Kavalan people, the Europeans relied heavily on the assistance of the indigenous Basay people from Keelung to coordinate and collect tributes in Yilan [@Kang2016]. This indirect approach allowed the maintenance of autonomy and flexibility of Kavalan people for negotiation tributes, such as deerskins or products, requested by the VOC [@Kang2016].

<!-- pre-exiting trade network -->
Before the arrival of the Europeans in northern Taiwan, an inter-insular trade network had been established since the early 15th century, involving Chinese merchants and indigenous groups in Tamsui and Keelung [@Chen2005]. Chinese traders transported goods such as textiles, porcelains, beads, and metal items to exchange for raw materials from Taiwan, such as gold, sulfur, and animal hides and meats from the local Basay people [@Chen2005; @Andrade2007]. This regional trade network was further connected to maritime trade between Ming China and the Ryukyu Kingdom during the 15th and 16th centuries and later to European long-distance trade in the 17th century [@Chen2005; @Von2019; @Wang2007].Yilan was included in this long-established trade network, where the Kavalan people exchanged products such as rice and deerskins for foreign goods and handicrafts from the Basay people [@Kang2003]. This is evidenced by the large amount of imported foreign goods found across archaeological sites in Yilan, indicating frequent trade activities, especially around the 17th century [@TaiwanHistory2005]. Yilan experienced frequent trading activities driven by the pursuit of foreign goods even before the Europeans arrival, and these activities continued during the European presence in northern Taiwan, likely with some occasional pressure from European powers. This unqiue historical and spatial contexts make Yilan an ideal location to study indigenous reactions to foreign influences and the associated trade networks from a long-term perspective in a pericolonial context [@Acabado2017; @Kang2016, pp. 31-78; @Trabert2017].

# Kiwulan site and previous studies
This study focuses on archaeological evidence from Kiwulan, a large multi-component site of with an upper component spanning from the late Iron Age to the historical periods (AD 1350-1850) [@Chen2007]. Kiwulan was inhabited by Kavalan people, as indicated by local oral history and the similarities between Kavalan material culture and archaeological artifacts [@Chen2004]. According to a mid-17th-century Dutch census, there were about 40 indigenous villages with approximately 10,000 people, and Kiwulan (referred to as Kibannoran in Dutch records) was the largest village, with 160 households and 711 people [@Nakamura1938]. In the early 19th century, the population of Kiwulan dramatically decreased to fewer than 100 people due to the population pressure of Han Chinese migrants, which led to the relocation of Kavalan people to eastern Taiwan [@Zhan1995].

The excavation at Kiwulan uncovered 262 units (4 x 4 m) and revealed abundant artifacts and features, such as middens, post-holes, and burials. These findings suggest a well-developed settlement with longhouses (approximately 40 x 80 m) situated side by side and burials located to the north of the residential area [@Chen2004]. In addition to locally made pottery, porcelains, stonewares, glass beads, carnelian beads and metal artifacts were largely found in the residential area and burials. Since there is no evidence of local production of beads or metallurgy, these goods are believed to have been acquired through a regional exchange network with people from outside the area.

# Previous research of

# Mateirals and methods
